API design for machine learning software: experiences from the scikit-learn project

### 2013-06-24 paper.tex

In \sklearn, we chose a representation of data that is as close as
possible to the matrix representation: datasets are encoded as NumPy
multidimensional arrays for dense data and as SciPy sparse matrices for sparse
data. While these may seem rather unsophisticated data representations when
compared to more object-oriented constructs, such as the ones used by
Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
on all NumPy and SciPy vector operations while keeping
the code short and readable.  This design choice has also been motivated by
the fact that, given their pervasiveness in many other scientific Python
packages, many scientific users of Python are already familiar with NumPy dense
arrays and SciPy sparse matrices.
From a practical point of view, these formats also provide a bunch of
data loading and conversion tools which make them very easy to use in many
contexts. Moreover, for tasks where the inputs are text files or semi-structured
objects, we provide \textit{vectorizer} objects that efficiently convert such
data to the NumPy or SciPy formats.

@@

clf = LogisticRegression(penalty="l1")
clf.fit(X_train, y_train)
\end{pythoncode}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{"l1"}, hence enabling
$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C}, which
defines the strength of the regularization) are not explicitly defined and are
thus set to the default values. Upon calling \texttt{fit}, the model is then
learned from the training arrays \texttt{X\_train} and \texttt{y\_train}.
(In this example, \texttt{fit} is called solely for its side effects.)
Since all estimators share the same interface, using a different learning algorithm is
as simple as replacing the constructor (the class name).
For example, to build a random forest on
the same training data, one would simply have to replace
\texttt{LogisticRegression(penalty="l1")} in the snippet above by \\
\texttt{RandomForestClassifier()}.

>>>>

In \sklearn, we chose a representation of data that is as close as
possible to the matrix representation: datasets are encoded as NumPy
multidimensional arrays for dense data and as SciPy sparse matrices for sparse
data. While these may seem rather unsophisticated data representations when
compared to more object-oriented constructs, such as the ones used by
Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
on efficient NumPy and SciPy vectorized operations while keeping
the code short and readable.  This design choice has also been motivated by
the fact that, given their pervasiveness in many other scientific Python
packages, many scientific users of Python are already familiar with NumPy dense
arrays and SciPy sparse matrices.
From a practical point of view, these formats also provide a bunch of
data loading and conversion tools which make them very easy to use in many
contexts. Moreover, for tasks where the inputs are text files or semi-structured
objects, we provide \textit{vectorizer} objects that efficiently convert such
data to the NumPy or SciPy formats.

@@

clf = LogisticRegression(penalty="l1")
clf.fit(X_train, y_train)
\end{pythoncode}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{"l1"}, hence enabling
$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C}, which
defines the strength of the regularization) are not explicitly defined and are
thus set to the default values. Upon calling \texttt{fit}, the model is then
learned from the training arrays \texttt{X\_train} and \texttt{y\_train},
and stored within the object for later use.
%% (In this example, \texttt{fit} is called solely for its side effects.)
Since all estimators share the same interface, using a different learning algorithm is
as simple as replacing the constructor (the class name).
For example, to build a random forest on
the same training data, one would simply have to replace
\texttt{LogisticRegression(penalty="l1")} in the snippet above by \\
\texttt{RandomForestClassifier()}.
